\documentclass[12pt]{letter}

\begin{document}


\hfill September 20,  2011
\newline

To Whom It May Concern-

I am writing to give my full support of Dr. Zhaoming Ma's application for
permanent residency in the United States of America.

Dr. Ma joined me as a Postdoctoral Researcher at Brookhaven National Laboratory
in the Fall of 2009.  Since that time we have worked on a new astronomical
instrument to study the phenomena of Dark Matter and Dark Energy in our
universe.  These phenomena are perhaps the greatest mysteries of our time, and
Dr. Ma has dedicated his career to their investigation.

I have a PhD in astrophysics from the University of Michigan, and am currently
an associate physicist at Brookhaven National Laboratory.  I have worked my
entire career on various experiments designed to illuminate the Dark Matter and
Dark Energy mysteries.  I brought Dr. Ma to Brookhaven to collaborate in the
study of these phenomena, and to help build the Dark Energy Survey, of which
I will say more below.

I will first say a bit about Dark Matter and Dark Energy, in order to clarify
my description of Dr. Ma's work.  Dark Matter is essentially a name for the
enormous amount of matter in our universe that is unseen.  We know it exists
because of its gravitational influence on our galaxy and other distant
galaxies.  Just as we know the mass of the Sun by studying the orbits of the
planets, we can infer the mass of galaxies from the orbits of stars in those
galaxies.  There is far more unseen mass than visible mass, and we call this
mass ``Dark''.  Similarly, Dark Energy is a name for a phenomenon, in this case
the inexplicable acceleration of our universe. We have known for almost a
century that the galaxies are all moving apart, but we expect gravity to slow
pull inward and slow them down.  Instead, some unknown energy is driving
galaxies in our universe apart at an every increasing rate.  Again we cannot
see the cause, so we name it ``Dark'' energy.

The Dark Energy Survey is a new experiment to study these phenomena.  We have
built a new camera, the largest CCD camera in the world, to survey the sky and
look for more evidence.  The camera is being installed as I write this letter,
and Dr. Ma and I will work throughout the winter and sprint to test this
camera, and prepare for survey operations next year.

Dr. Ma's work on the Dark Energy Survey has fallen into two broad categories:
so-called ``gravitational lensing'' measurements and survey design, which I
will describe in turn.  

Dr. Ma has spent most of his time on gravitational lensing measurements.
Gravitational lensing is a beautiful phenomenon where the light from distant
background galaxies gets ``lensed'' by closer foreground galaxies as it makes
its way to us.  Much as the image of a word in a book is enlarged and distorted
by a magnifying glass, the image of a background galaxy is magnified and
distorted by the presence of another foreground galaxy.  This magnification and
distortion occurs because the light is deflected as it passes through its
gravity well of the foreground galaxy.  And much like the magnifying glass, the
amount of distortion depends on how strong the lens is and how close or far it
is from your eye.  In gravitational lensing, the ``strength'' of the lens is
just the mass of the foreground galaxy.

Gravitational lensing is the most powerful phenomena for the study of Dark
Matter and Dark Energy, because both phenomena effect lensing.  Because the
lens ``strength'' is just mass, lensing is used to count the Dark Matter in
galaxies.  Similarly, because the lens effect depends on how far away the lens
is, we can use the inferred lens magnification or distortion to tell us how
far apart galaxies are over the history of the universe, and thus see how much
Dark Energy has driven galaxies apart.

The difficulty is that the telescope has its own astigmatism, and this also
distorts images.  This effect must be understood if we are to learn about Dark
Energy.  To make matters worse, it turns out there is not enough information in
an image to determine this effect.  But Dr. Ma has developed the best method in
the world to deal with this lack of information.  He characterizes the expected
patterns by studying all the data from a survey, which reduces greatly the
range of possibilities, and makes determination of the effect possible.  This
is similar to how the best analysts study the stock market; only by studying
the entire history can one understand why the market acts as it does on a
particular day.  Dr. Ma's method is crucial to the success of the survey, and
he will employ it continuously as the survey progresses to ensure that we can
learn as much as possible about Dark Energy and Dark Matter.

Dr. Ma also studied which survey design would give the most useful information
about Dark Energy.  For example, he calculates the trade-off between exposing
images for a very long time or taking many more shorter exposures; how best to
do take color images as opposed to monochrome; and how and when to look at
certain parts of the sky as opposed to others.  These decisions critically
impact the quality of the survey, and Dr. Ma's contribution is recognized
within the Dark Energy Survey as crucial for the project's success.


{\noindent Erin Sheldon}
\newline
{\noindent Brookhaven National Laboratory}


\end{document}
